Towards an Automated Sportscasting System for Chess

### 2011-05-18 paper/paper.tex

\subsection{Manually-generated features}
\begin{table*}
\centering
\begin{tabular}{lp{0.6\textwidth}}
\hline
\textbf{Category} & \textbf{Description} \\ \hline
Position & Piece location, potential moves per piece, material count, square control. \\
Previous move & Capturing move, recapturing move, en passant move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
\hline
\end{tabular} 
\caption{List of chess-specific manual features per position.}
\label{table:features}
\end{table*}

@@

A chess-specific approach would be able to perfectly identify the more precisely-described tags like {\tt doubleAttack} and {\tt queenCapture} by analyzing the future branches and looking for specific outcomes. In addition, it would be able to identify less concrete tags like {\tt clearance} and {\tt kingToBadSquares}. However, one advantage of the statistical approach is that feature functions can be written once and reused for all inputs. In contrast, in order to effectively identify tags using a chess engine, each tag must have a specific heuristic function written for that express purpose to identify certain properties in future states. A better automated sportscasting for chess will leverage the strengths of each approach by using the chess-specific features when they are available, and combining that with results from statistical learning. For instance, while the chess engine can easily identify imminent mate combinations, it would require a coordination of both the engine and the statistical learning tool to identify threshold functions like {\tt kingToBadSquares}. Machine learning would be used to optimize weights and thresholds (e.g. for whether a king’s position is bad) while the engine would be used to explore future paths.

\section{Future Work}
There are three areas in which future work can make progress on automated sportscasting. It could improve the performance of the auto-encoder so that it can be trained on a larger data set. Furthermore, it could focus on trying different methods to improve the classification quality by leveraging chess-specific tools and features like chess engines and evaluation tools. Moreover, it could investigate an orthogonal path by applying our general approach to other games, including complex ones like StarCraft and simpler ones like Checkers.

@@

>>>>

\subsection{Manually-generated features}
\begin{table*}
\centering
\begin{tabular}{lp{0.6\textwidth}}
\hline
\textbf{Category} & \textbf{Description} \\ \hline
Position & Piece location, potential moves per piece, material count, square control. \\
Previous move & Capturing move, recapturing move, \emph{en passant} move, long castling move, short castling move, check move, stalemate move, checkmate move, promotion move. \\
\hline
\end{tabular} 
\caption{List of chess-specific manual features per position.}
\label{table:features}
\end{table*}

@@

\section{Future Work}
We identify three areas in which future work can make progress on automated sportscasting for chess. It could improve the performance of the auto-encoder so that it can be trained on a larger data set. Furthermore, it could focus on trying different methods to improve the classification quality by leveraging chess-specific tools and features like chess engines and evaluation tools. Moreover, it could investigate an orthogonal path by applying our general approach to other games, including complex ones like StarCraft and simpler ones like Checkers.

@@

A chess-specific rule-based approach should be able to easily identify the more precisely-described themes like {\tt doubleAttack} and {\tt queenCapture} by analyzing the future branches of the game and looking for specific outcomes. However, one advantage of the statistical approach is that feature functions can be written once and reused for all possible themes. In addition, it would help in identifying more subjective themes like {\tt clearance} and {\tt kingToBadSquares} based more on human judgement. In contrast, in order to effectively identify tags using a chess engine, each theme must have a specific heuristic function written for that express purpose to identify certain properties in future states. A better automated sportscasting for chess will leverage the strengths of each approach by using the chess-specific features when they are available, and combining that with results from statistical learning. For instance, while the chess engine can easily identify imminent mate combinations, it would require a coordination of both the engine and the statistical learning tool to identify threshold functions like {\tt kingToBadSquares}. Machine learning would be used to optimize weights and thresholds (e.g. for whether a king’s position is bad) while the engine would be used to explore future paths.
